{ \fontfamily{ptm}\selectfont Catalyst Meta Acceleration Framework: The history and the gist of it }

### 2024-11-19 Notebook LaTEX/Catalyst Accelerations/main.tex
>w<, changes

        The same inequality is called ``Fundamental Proximal Gradient Inequality'' in Amir Beck's book \cite{beck_first-order_2017}, Theorem 10.16. 

@@

        \right) &\le \frac{\epsilon_k}{k}

@@

@@

        This comment has upmost practical importance because it tells us how to bound the error $\epsilon_k$ to achieve accelerated convergence rate. 
        In theory, $\epsilon_k$ decreases at a rate related to $F(x_0) - F^*$. 
        It requires some knowledge about $F^*$ in prior. 
        Therefore, controlling $\epsilon_k$ is still elusive in general in a practical context. 
        To see how the error is controlled for the inexact proximal point evaluation, we refer the readers to Lemma B.1 in Lin's 2015 paper \cite{lin_universal_2015}. 
    \end{remark}
    For theoretical interests, there is a major difference between Lin's approach and Guler's approach. 
    Lin didn't formulate any of the intractable quantities in the definitions for his Nesterov's estimating sequence $\phi_k$. 
    One major innovation is Lemma A.7 in Lin's 2015 paper \cite{lin_universal_2015}. 
    The lemma allows the analysis Nesterov's estimating sequence to be carried through without using intractable quantities: $\mathcal M^{\kappa^{-1}}(\mathcal J_{k^{-1}}y_{k - 1}, y_{k - 1}), \mathcal J_{\kappa^{-1}}y_{k - 1}$. 
    \begin{lemma}[Lin's inexact proximal inequality]
        Let $F$ be a $\mu\ge 0$ strongly convex.
        Suppose $x_k$ is an inexact proximal point evaluation of $x_k \approx \mathcal J_{\kappa^{-1}} y_{k - 1}$ with $\kappa$ fixed. 
        Assume the approximation error is characterized by $\mathcal M^{\kappa^{-1}}(x_k; y_{k - 1}) - \mathcal M^{\kappa^{-1}}(\mathcal J_{\kappa^{-1}} y_{k - 1}, y_{k - 1}) \le \epsilon_k$. 
        Denote $x_k^* = \mathcal J_{\kappa^{-1}} y_{k - 1}$ to be the exact evaluation of the proximal point then for all $x$: 
        \begin{align*}
            F(x) &\ge 
            F(x_k) + \kappa \langle y_{k - 1} - x_k, x - x_k\rangle
            + \frac{\mu}{2}\Vert x - x_k\Vert^2 
            + (\kappa + \mu)\langle  x_k - x_k^*, x - x_k\rangle 
            - \epsilon_k.
        \end{align*}
    \end{lemma}
    \begin{remark}
        The lemma plays a key role because it allows Lin to denote his Nesterov's estimating sequence to be for all $k \ge0$: 
        \begin{align*}
            \phi_k(x) := 
            (1 - \alpha_{k - 1})\phi_{k - 1}(x)
            + \alpha_{k - 1}\left(
                F(x_k) + \kappa\langle y_{k - 1} - x_k, x - x_k \rangle
                + \frac{\mu}{2}\Vert x - x_k\Vert^2
            \right). 
        \end{align*}
        It is void of intractable quantities. 
    \end{remark}

@@

        Next move is to simplify the term $\Vert v_{k + 1} - y_k\Vert^2$. 
        We are doing this in advance so that later on the implicit descent conditions $F(x_{k + 1}) \le \phi_k^*$ will have easier algebra. 

@@

                \alpha_k(1 - \alpha_k)v_k - \alpha_k g_k 

@@

                    \gamma_{k + 1} - \mu \alpha_k &= (1 - \alpha_k)\gamma_k

@@

                \alpha_k(1 - \alpha_k)v_k - \alpha_k g_k 

@@

                \alpha_k(1 - \alpha_k)(v_k - y_k) 

@@

        Taking the norm of that we have: 

@@

@@

    \subsection{Exact accelerated PPM}

@@

@@

    \subsection{Inexact accelerated PPM}
        This section proofs the lemma that characterizes the inexact PPM errors for Guler's accelerated inexact PPM. 
        Now, we prove 
        \hyperref[thm:guler-inexact-ppm-bound]{Theorem \ref*{thm:guler-inexact-ppm-bound}}.

        The upper bound is proved.

\section{Proofs for Catalyst Meta Acceleration}
    In this section, we prove the inexact proximal inequality and give the Nesterov's estimating sequence void of intractable quantities. 
    The notation follows from 
    \hyperref[sec:lin-2015]{Section \ref*{sec:lin-2015}}. 
    Let's now prove the inexact proximal inequality. 
    Recall inexact evaluation $x_k \approx \mathcal J_{\kappa^{-1}}y_{k - 1}$ such that $\mathcal M^{\kappa^{-1}}(x_k; y_{k - 1}) - \mathcal M^{\kappa^{-1}}(\mathcal J_{\kappa^{-1}}y_{k - 1}; y_{k - 1}) \le \epsilon_k$; $x_k^* = \mathcal J_{\kappa^{-1}}y_{k - 1}$ to be the exact evaluation.

\section{Proofs for 4WD Catalyst Acceleration}
    We prove
    \hyperref[thm:basic-4wd-catalyst]{Theorem \ref*{thm:basic-4wd-catalyst}}
    in this section. 
    The lemma below characterize the lower and upper bound on the $(\alpha_k)_{k \ge 0}$ which ultimately control the convergence rate. 
    \begin{lemma}{app:lemma:momentum-sequence-bounds}

>>>>

@@

        \right) &\le \frac{\epsilon_k}{\lambda_k}

@@

        The inequalities are from PL-condition and strong convexity. 
        They are now standard results in convex optimizations. 
        For a proof of the theorem, see 
        \hyperref[app:sec:inxt-acc-ppm]{Appendix \ref*{app:sec:inxt-acc-ppm}}. 

@@

        This comment has upmost practical importance because it tells us how to bound the error $\epsilon_k$ to achieve accelerated convergence rate. 
        In theory, $\epsilon_k$ decreases at a rate related to $F(x_0) - F^*$. 
        It requires some knowledge about $F^*$ in prior. 
        Therefore, controlling $\epsilon_k$ is still elusive in general in a practical context. 
        To see how the error is controlled for the inexact proximal point evaluation, we refer the readers to Lemma B.1 in Lin's 2015 paper \cite{lin_universal_2015}. 
    \end{remark}
    For theoretical interests, there is a major difference between Lin's approach and Guler's approach. 
    Lin didn't formulate any of the intractable quantities in the definitions for his Nesterov's estimating sequence $\phi_k$. 
    One major innovation is Lemma A.7 in Lin's 2015 paper \cite{lin_universal_2015}. 
    The lemma allows the analysis Nesterov's estimating sequence to be carried through without using intractable quantities: $\mathcal M^{\kappa^{-1}}(\mathcal J_{k^{-1}}y_{k - 1}, y_{k - 1}), \mathcal J_{\kappa^{-1}}y_{k - 1}$. 
    \begin{lemma}[Lin's inexact proximal inequality]\label{lemma:lin-ixct-prox-ineq}
        Let $F$ be a $\mu\ge 0$ strongly convex and fix $\kappa$. 
        If $x_k$ is an inexact proximal point evaluation of $x_k \approx \mathcal J_{\kappa^{-1}} y_{k - 1}$ such that there exists $\epsilon_k$ where $\mathcal M^{\kappa^{-1}}(x_k; y_{k - 1}) - \mathcal M^{\kappa^{-1}}(\mathcal J_{\kappa^{-1}} y_{k - 1}, y_{k - 1}) \le \epsilon_k$. 
        Denote $x_k^* = \mathcal J_{\kappa^{-1}} y_{k - 1}$ it has for all $x$: 
        \begin{align*}
            F(x) &\ge 
            F(x_k) + \kappa \langle y_{k - 1} - x_k, x - x_k\rangle
            + \frac{\mu}{2}\Vert x - x_k\Vert^2 
            + (\kappa + \mu)\langle  x_k - x_k^*, x - x_k\rangle 
            - \epsilon_k.
        \end{align*}
    \end{lemma}
    \begin{remark}
        The lemma plays a key role because it allows Lin to denote his Nesterov's estimating sequence to be for all $k \ge0$: 
        \begin{align*}
            \phi_0 &:= F(x_0) + \frac{\gamma_0}{2}\Vert x - v_0\Vert^2, 
            \\
            \phi_k(x) &:= 
            (1 - \alpha_{k - 1})\phi_{k - 1}(x)
            + \alpha_{k - 1}\left(
                F(x_k) + \kappa\langle y_{k - 1} - x_k, x - x_k \rangle
                + \frac{\mu}{2}\Vert x - x_k\Vert^2
            \right). 
        \end{align*}
        It is void of intractable quantities. 
    \end{remark}

@@

        Next move from Nesterov was to simplify the term $\Vert v_{k + 1} - y_k\Vert^2$. 

@@

                \gamma_k(1 - \alpha_k)v_k - \alpha_k g_k 

@@

                    -(1 - \alpha_k)\gamma_k
                    &= - \gamma_{k + 1} + \mu \alpha_k

@@

                \gamma_k(1 - \alpha_k)v_k - \alpha_k g_k - 

@@

                \gamma_k(1 - \alpha_k)(v_k - y_k) 

@@

        Next we have

@@

        The goal of having $\phi_k^*$ represented by iterates $v_k, y_k, g_k$ is because it assists in the future when an inductive hypothesis $\phi_k^* \ge f(x_k) \ge l_F(x_k; y_k) + \frac{\mu}{2}\Vert x - y_k\Vert^2$ is made. 
        We did things in advance here, nothing more. 

@@

    \subsection{Exact accelerated PPM}\label{sec:app:exct-acc-ppm}

@@

            In here, we substitute the inductive hypothesis $\phi_k^* \ge f(x_k)$ and skipped the canonical representation of $\phi_k^*$.

            A more critical observation here is the Nesterov's estimating sequence from 
            \hyperref[app:sec:thm-claim-acc-prox-grad]{Section \ref*{app:sec:thm-claim-acc-prox-grad}} is the same as Guler's Accelerated PPM if we choose $L^{-1} = \lambda_k$,$\mu = 0$ and $g \equiv 0$ because: 
            \begin{align*}
                l_F(x; y_k) 
                &= 
                F\left(\widetilde{\mathcal J}_{L^{-1}} y_k\right) + 
                \left\langle 
                    L\left(y_k - \widetilde{\mathcal J}_{L^{-1}}y_k\right), 
                    x - y_k
                \right\rangle
                + 
                \frac{1}{2L} 
                \left\Vert 
                    L \left(y_k - \widetilde{\mathcal J}_{L^{-1}}y_k\right)
                \right\Vert^2
                \\
                &= 
                F\left(\widetilde{\mathcal J}_{L^{-1}} y_k\right) + 
                \left\langle 
                    L\left(y_k - \widetilde{\mathcal J}_{L^{-1}}y_k\right), 
                    x - \widetilde{\mathcal J}_{L^{-1}}y_k
                \right\rangle
            \end{align*}
            When $g\equiv 0$, $L^{-1} = \lambda_k$, the proximal gradient operator has $\widetilde{\mathcal J}_{L^{-1}} = \mathcal J_{\lambda_k}(y_k)$, making $\phi_k(x)$ here the same as accelerated proximal gradient. 
        \end{remark}
    \subsection{Inexact accelerated PPM}\label{app:sec:inxt-acc-ppm}
        This section proofs the lemma that characterizes the inexact PPM errors for Guler's accelerated inexact PPM. 
        Now, we prove 
        \hyperref[thm:guler-inexact-ppm-bound]{Theorem \ref*{thm:guler-inexact-ppm-bound}}.

        The upper bound is proved. 
        \begin{remark}
            This is nothing new. 
            First inequality is a direct consequence of strong convexity and the second inequality is the PL-inequality implied by strong convexity. 
        \end{remark}

\section{Proofs for Catalyst Meta Acceleration}
    In this section, we prove the inexact proximal inequality and give the Nesterov's estimating sequence void of intractable quantities. 
    The notation is the same as
    \hyperref[sec:lin-2015]{Section \ref*{sec:lin-2015}}. 
    Recall inexact evaluation $x_k \approx \mathcal J_{\kappa^{-1}}y_{k - 1}$ such that $\mathcal M^{\kappa^{-1}}(x_k; y_{k - 1}) - \mathcal M^{\kappa^{-1}}(\mathcal J_{\kappa^{-1}}y_{k - 1}; y_{k - 1}) \le \epsilon_k$; $x_k^* = \mathcal J_{\kappa^{-1}}y_{k - 1}$ to be the exact evaluation.

    \subsection{Inexact proximal inequality}
            This is the proof for 
            \hyperref[lemma:lin-ixct-prox-ineq]{Lemma \ref*{lemma:lin-ixct-prox-ineq}}. 
            Define $G_k := \mathcal M_F^{\kappa^{-1}}(\cdot, y_{k - 1}), G_k^* = \mathcal M_F^{\kappa^{-1}}(x_k^*, y_{k - 1})$. 
        $G_k$ is a $\mu + \kappa$ convex function which implies quadratic growth over minimizer $x_k^*$: 
            (\forall x)\quad G_k(x) &\ge G_k^* + \frac{\kappa + \mu}{2}\Vert x - x_k^*\Vert^2. 
        Substituting definitions:
            F(x) &\ge 
            G_k(x_k) + (G_k^* - G_k(x_k)) + 
            \frac{\mu + \kappa}{2}\Vert x - x_k^*\Vert^2 
            - \frac{\kappa}{2}\Vert x - y_{k - 1}\Vert^2
            G(x_k) - \epsilon_k + \frac{\kappa + \mu}{2}\Vert x - x_k^*\Vert^2
            - \frac{\kappa}{2}\Vert x - y_{k - 1}\Vert^2
            &= G_k(x_k) - \epsilon_k + \frac{\kappa + \mu}{2}
                \Vert x - x_k - x_k + x_k^*\Vert^2
            - \frac{\kappa}{2}\Vert x - y_{k - 1}\Vert^2
            &= G_k(x_k) - \epsilon_k + \frac{\kappa + \mu}{2}
                \Vert x - x_k\Vert^2 + \Vert x_k - x_k^*\Vert^2
                2\langle x - x_k, x_k - x_k^*\rangle
            - \frac{\kappa}{2}\Vert x - y_{k - 1}\Vert^2
                G_k(x_k) + \frac{\kappa}{2}\Vert x - x_k\Vert^2 - \frac{\kappa}{2}\Vert x - y_{k - 1}\Vert^2
            - \epsilon_k
            \\ &\quad 
                + \frac{\mu}{2}\Vert x - x_k\Vert^2
                + \frac{\kappa + \mu}{2}\Vert x_k - x_k^*\Vert^2
                + (\kappa + \mu)\langle x - x_k, x_k - x_k^*\rangle. 
        Simplify terms inside the parenthesis: 
            G_k(x_k) 
            + \frac{\kappa}{2}\Vert x - x_k\Vert^2 - \frac{\kappa}{2}\Vert x - y_{k - 1}\Vert^2
            F(x_k) + \frac{\kappa}{2}\Vert x_k - y_{k - 1}\Vert^2
            + \frac{\kappa}{2}\Vert x - x_k\Vert^2 - \frac{\kappa}{2}\Vert x - y_{k - 1}\Vert^2
            F(x_k) + 
            \frac{\kappa}{2}\left(
                \Vert x_k - y_{k - 1}\Vert^2 - \Vert x - y_{k - 1}\Vert^2
            + \frac{\kappa}{2} \Vert x - x_k\Vert^2
            F(x_k) + 
            \frac{\kappa}{2}\left(
                \Vert x_k - x\Vert^2 
                2\langle x_k - x, x - y_{k - 1} \rangle
            \frac{\kappa}{2} \Vert x - x_k\Vert^2
            F(x_k) + 
            \kappa\Vert x_k - x\Vert^2
            \kappa\langle x_k - x, x - y_{k - 1}\rangle
            F(x_k) + 
            \kappa\langle x_k - x, x - y_{k - 1} + x_k - x\rangle
            &= F(x_k) + \kappa\langle x_k - x, x_k - y_{k - 1} \rangle. 
        Therefore, it has the inequality: 
            F(x) 
            F(x_k) + \kappa\langle x_k - x, x_k - y_{k - 1} \rangle
            - \epsilon_k
                + \frac{\mu}{2}\Vert x - x_k\Vert^2
                + \frac{\kappa + \mu}{2}\Vert x_k - x_k^*\Vert^2
                + (\kappa + \mu)\langle x - x_k, x_k - x_k^*\rangle
            & \ge
            F(x_k) + \kappa\langle x_k - x, x_k - y_{k - 1} \rangle
            - \epsilon_k
            + \frac{\mu}{2}\Vert x - x_k\Vert^2
            + (\kappa + \mu)\langle x - x_k, x_k - x_k^*\rangle. 
        Re-arranging it to a form by the writer's preference:
            F(x) - F(x_k) - \kappa\langle x_k - x, x_k - y_{k - 1} \rangle
            - \frac{\mu}{2}\Vert x - x_k\Vert^2
            - \epsilon_k
            + (\kappa + \mu)\langle x - x_k, x_k - x_k^*\rangle. 
        Now we have the RHS to be exclusively about the inexact evaluation $x_k \approx \mathcal J_{\kappa^{-1}} y_{k - 1}$, and $G_k(x_k) - G^*_k \le \epsilon_k$.

        \end{proof}
        \begin{remark}

@@

\section{Proofs for 4WD Catalyst Acceleration}
    We prove
    \hyperref[thm:basic-4wd-catalyst]{Theorem \ref*{thm:basic-4wd-catalyst}}
    in this section. 
    The lemma below characterize the lower and upper bound on the $(\alpha_k)_{k \ge 0}$ which ultimately control the convergence rate. 
    \begin{lemma}\label{app:lemma:momentum-sequence-bounds}
